---
title: "Data Structures in R"
author: "Elizabeth Permina"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DiagrammeR)
```

## Data classes

Everything is a vector
Character (string), numeric, integer, double, factor
Factor is an integer
## Data types

```{r}
grViz("
digraph boxes_and_circles {

  # a 'graph' statement
  graph [overlap = true, fontsize = 10]

  # several 'node' statements
  node [shape = box,
        fontname = Helvetica]
  DataStructures; homogenous; heterogenous; atomicVector; matrix; list; dataFrame; dataframe; tbl_df; tbl



  # several 'edge' statements
  DataStructures->homogenous
  DataStructures->heterogenous
  homogenous->atomicVector homogenous->matrix
  heterogenous-> dataFrame
  heterogenous-> list
  dataFrame -> dataframe
  dataFrame -> tbl_df
  dataFrame -> tbl
  
}
")
```

Atomic vector, assumed data classes

```{r vector}
n_vector<-c(1,2,3,5) 
n_vector
another_vector<-c(rep(1, 3), rep(2, 5), runif(4, min = 3, max = 6))
another_vector
ch_vector<-c("apple", 45, "hot cross bun")
ch_vector
c(another_vector, ch_vector)

```

A matrix is the special case of a two-dimensional array.

A matrix example:

```{r matrix}
m<- matrix(c(1,2,3, 11,12,13), nrow = 2, ncol = 3, byrow = TRUE,
               dimnames = list(c("row1", "row2"),
                               c("C.1", "C.2", "C.3")))
m
```

A list example

```{r}
my_list <- list(a=ch_vector, b=n_vector, c = another_vector, m = m)
my_list
```

```{r}
my_list[1]
```

```{r}
my_list[[1]]
```

```{r}
my_list$a[[1]]
```
you can even do a list of lists

## Working with data frames
Inconsistent data type
### Subsetting in base R
list of column names, boolean vector, list of column numbers
[ and [[
$ operator

#### dropping levels

### tibble - tbl and tbl_df
#### vector or dataframe
